---
title: "Partical machine learning - Course project"
author: "Houmin Zhong"
date: "December 3, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Partical machine learning - Course project Nov 2016
## Instruction
Data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.
They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
The goal of your project is to predict the manner in which they did the exercise. 
This is the "classe" variable in the training set.

Describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

### Data processing 
```{r, echo =TRUE}
training <- read.csv('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv', header=T) # read the training data
testing <- read.csv('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv', header=T) # read the testing data
table(training$classe)

trainingCleann<-training[training$new_window=="no",] # using rows when new_window is no. 
column.has.na <- apply(trainingCleann, 2, function(x){any(is.na(x))|any(x=='')}) # get the column numbers which have no empty data. 

trainingCleann<-trainingCleann[,!column.has.na] #columns have no empty data. 
trainingCleann<-trainingCleann[,-c(1,3,4,5,6)] # delete the first, the timestamp, and the new_window colums. Assuming they are not important for prediction. 
testingCleann<-testing[,!column.has.na] #columns have no empty data. 
testingCleann<-testingCleann[,-c(1,3,4,5,6)] # delete the first, the timestamp, and the new_window colums. Assuming they are not important for prediction. 
```

### Built models
```{r, echo =TRUE}
library(caret)
library(ggplot2)
set.seed(3721)
# three models would be built
# predicting with trees
rpartFit<- train(classe~.,method="rpart", data=trainingCleann)
# Linear Discriminant Analysis
ldaFit <- train(classe~.,method="lda", data=trainingCleann) 
# quadratic Discriminant Analysis 
qdaFit <- train(classe~.,method="qda", data=trainingCleann) 
```

### Finding accuracy
```{r, echo =TRUE}
rpartFit$results
qdaFit$results
ldaFit$results

print(paste0("qdaFit has the highest accuracy among three models which is ",round(qdaFit$results[2],2), ". Therefore, we choose the Quadratic Discriminant Analysis model as the final prediction model." ))
```

### Test set data participants movement classe prediction 
```{r figure.width =7, fig.height=6, echo =TRUE}
test_qda<- predict(qdaFit, testingCleann) # using the Quadratic Discriminant Analysis model to predict participants movement classe in the test dataset.
Participant<-testingCleann$user_name
# plot predicted movement classe for each problem_id in test set.
qplot(testingCleann$problem_id,test_qda, colour=Participant, main="Test Set Movement Predictions", xlab = "Problem id", ylab = "Movement classe") 
```

### Thank you for checking my work. 