---
title: "Partical machine learning - Course project"
author: "Houmin Zhong"
date: "December 17, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Partical machine learning - Course project Nov 2016
## Instruction
Data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.
They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
The goal of your project is to predict the manner in which they did the exercise. 
This is the "classe" variable in the training set.

Describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

### Data processing 
Data are downloaded and cleaned.
```{r, echo =TRUE}
require(caret)
training <- read.csv('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv', header=T, na.strings=c("NA", "#DIV/0!","")) # read the training data
testing <- read.csv('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv', header=T, na.strings=c("NA", "#DIV/0!","")) # read the testing data
table(training$classe)

trainingCleann<-training[training$new_window=="no",] # using rows when new_window is no. 
column.has.na <- apply(trainingCleann, 2, function(x){any(is.na(x))|any(x=='')}) # get the column numbers which have no empty data. 

trainingCleann<-trainingCleann[,!column.has.na] #columns have no empty data. 
trainingCleann<-trainingCleann[,-c(1,3,4,5,6)] # delete the first, the timestamp, and the new_window colums. Assuming they are not important for prediction. 
rm(training)
testingCleann<-testing[,!column.has.na] #columns have no empty data. 
testingCleann<-testingCleann[,-c(1,3,4,5,6)] # delete the first, the timestamp, and the new_window colums. Assuming they are not important for prediction. 
rm(testing)

```


### Data Partitioning
Training data set is splited into train (80%) and test (20%) data set for cross validation
```{r}
set.seed(123456)
inTrain<- createDataPartition(y = trainingCleann$classe, p = 0.8, list = FALSE) # create 80% and 20% partitions on the training data
train <- trainingCleann[inTrain,]# train data set
test <- trainingCleann[-inTrain,]# test data set
rm(trainingCleann)
```


### Built models
Test data would be predicted with trees, Linear Discriminant Analysis model, and quadratic Discriminant Analysis model.
```{r, echo =TRUE}
set.seed(3721)
# three models would be built
# predicting with trees
rpartFit<- train(classe~.,method="rpart", data=train)
# Linear Discriminant Analysis
ldaFit <- train(classe~.,method="lda", data=train) 
# quadratic Discriminant Analysis 
qdaFit <- train(classe~.,method="qda", data=train) 
```

### Finding accuracy using Cross Validation
```{r, echo =TRUE}
rpartFit$results
qdaFit$results
ldaFit$results
test_rpart<- predict(rpartFit, test) # predicet the test data using rpartFit model
test_qda<- predict(qdaFit, test) # predicet the test data using qdaFit model
test_lda<- predict(qdaFit, test) # predicet the test data using ldaFit model 

rpart_CM<-confusionMatrix(test_rpart, test$classe) # confusion Matrix of tree model
rpart_CM
qda_CM<-confusionMatrix(test_qda, test$classe) # confusion Matrix of qda model
qda_CM
lda_CM<-confusionMatrix(test_lda, test$classe) # confusion Matrix of lda model
lda_CM

print(paste0("qdaFit has the highest accuracy among three models which is ",round(qda_CM$overall[1],2), ". Therefore, we choose the Quadratic Discriminant Analysis model as the final prediction model." ))
```

### Validation
Final Test (validation) data set participants movement classe (problem id) prediction. 
```{r figure.width =7, fig.height=6, echo =TRUE}
finalTest_qda<- predict(qdaFit, testingCleann) # using the Quadratic Discriminant Analysis model to predict participants movement classe in the testingCleann dataset.
Participant<-testingCleann$user_name
# plot predicted movement classe for each problem_id in testingCleann set.
qplot(testingCleann$problem_id,finalTest_qda, colour=Participant, main="Movement Predictions", xlab = "Problem id", ylab = "Movement classe") 
```

### Thank you for checking my work. 
